---
title: "Microbial Response to Intermittent Permafrost Thaw -- Clustering"
author: 'Authors: [Karl J. Romanowicz](https://kromanowicz.github.io/) and George W. Kling'
output:
  html_notebook:
    theme: spacelab
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  html_document:
    toc: yes
    toc_depth: '5'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '5'
---

__________________________________________________

**R Notebook:** <font color="green">Provides reproducible analysis for **Hierarchical Clustering** data in the following manuscript:</font>

**Citation:** Romanowicz KJ and Kling GW. (***In Press***) Summer thaw duration is a strong predictor of the soil microbiome and its response to permafrost thaw in arctic tundra. ***Environmental Microbiology***. [https://doi.org/10.1111/1462-2920.16218](https://doi.org/10.1111/1462-2920.16218)

**GitHub Repository:** [https://github.com/kromanowicz/2022-Annual-Thaw-Microbes](https://github.com/kromanowicz/2022-Annual-Thaw-Microbes)

**NCBI BioProject:** [https://www.ncbi.nlm.nih.gov/bioproject/?term=PRJNA794857](https://www.ncbi.nlm.nih.gov/bioproject/?term=PRJNA794857)

**Accepted for Publication:** <font color="green">22 September 2022</font> *Environmental Microbiology*

# Experiment

This R Notebook provides complete reproducibility of the data analysis presented in ***"Summer thaw duration is a strong predictor of the soil microbiome and its response to permafrost thaw in arctic tundra"*** by Romanowicz and Kling.

<font color="green">This pipeline uses amplicon sequence variants (ASVs) generated from 16S rRNA gene sequences in hierarchical clustering analysis to determine statistically significant clusters based on soil depth and shared taxonomy.</font>

```{r setup, include=FALSE}
# Set global options for notebook
knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/TundraPro18"))
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, results='hide', warning=FALSE}
# Make a vector of required packages
required.packages <- c("corrr","data.table","devtools","dplyr","forcats","ggalluvial","ggdendro","ggplot2","ggpubr","grid","gridExtra","knitr","magrittr","microeco","patchwork","pheatmap","pvclust","qiime2R","RColorBrewer","tidyr","UpSetR","vegan")

# Load required packages
lapply(required.packages, library, character.only = TRUE)
```

```{r include=FALSE}
# set.seed is used to fix the random number generation to make the results repeatable
set.seed(123)
```

# Hierarchical Clustering

This analysis includes heatmaps of z-scored relative abundance taxonomic data and hierarchical clustering by soil depth and taxa using pvclust and 10,000 bootstrap iterations.

## Toolik MAT

Import mean abundance of taxa by depth for TTT heatmap
```{r}
asv.taxa.ttt.mean <- read.csv("QIIME/SILVA/R_Data/taxa.mean.ttt.csv")

colnames(asv.taxa.ttt.mean)<-c("Phylum","0-10","10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90")

# Convert the first column (Phylum) into rownames
rownames(asv.taxa.ttt.mean) <- asv.taxa.ttt.mean$Phylum
asv.taxa.ttt.mean<-as.data.frame(asv.taxa.ttt.mean[-1])

# Convert dataframe into a matrix for heatmap
asv.taxa.ttt.mean<-as.matrix(asv.taxa.ttt.mean)

# Scale matrix values to generate Z-scores
asv.taxa.ttt.mean<-scale(t(asv.taxa.ttt.mean))

# Specify RColorBrewer custom color palette
col <- colorRampPalette(brewer.pal(10, "RdYlBu"))(256)
```

```{r}
# Pheatmap
asv.taxa.ttt.mean.pheatmap <- pheatmap(asv.taxa.ttt.mean, clustering_method = "average", cluster_cols = TRUE, cluster_rows = TRUE, cutree_cols = 3, cutree_rows = 3, angle_col=45, fontsize_col=8, legend = FALSE)

asv.taxa.ttt.mean.pheatmap.v2 <- pheatmap(asv.taxa.ttt.mean, clustering_method = "average", cluster_cols = TRUE, cluster_rows = FALSE, cutree_cols = 3, gaps_row = c(2, 4), angle_col=45, fontsize_col=10, legend = FALSE, cellheight=16, cellwidth=18)

#Export as .eps (width:650, height:600; "ttt.heat.silva.eps")
```

### Soil Depth pvclust

TTT Soil Depth pvclust Analysis
```{r}
asv.taxa.ttt.mean.pvclust <- read.csv("QIIME/SILVA/R_Data/taxa.mean.ttt.csv")

colnames(asv.taxa.ttt.mean.pvclust)<-c("Phylum","0-10","10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90")

# Convert the first column (Phylum) into rownames
rownames(asv.taxa.ttt.mean.pvclust) <- asv.taxa.ttt.mean.pvclust$Phylum
asv.taxa.ttt.mean.pvclust<-as.data.frame(asv.taxa.ttt.mean.pvclust[-1])

# Scale data
asv.taxa.ttt.mean.pvclust<-scale(asv.taxa.ttt.mean.pvclust)
```

```{r}
result.10k.ttt <- pvclust(asv.taxa.ttt.mean.pvclust, method.dist="cor", method.hclust="average", nboot=10000, parallel=TRUE)

plot(result.10k.ttt)
pvrect(result.10k.ttt, alpha=0.95)
```

### Taxonomy pvclust

TTT Soil Taxa pvclust Analysis
```{r}
asv.depth.ttt.mean.pvclust <- read.csv("QIIME/SILVA/R_Data/taxa.mean.ttt.cols.csv")

# Convert the first column (Phylum) into rownames
rownames(asv.depth.ttt.mean.pvclust) <- asv.depth.ttt.mean.pvclust$Phylum
asv.depth.ttt.mean.pvclust<-as.data.frame(asv.depth.ttt.mean.pvclust[-1])

# Scale data
asv.depth.ttt.mean.pvclust<-scale(asv.depth.ttt.mean.pvclust)
```

```{r}
result.10k.ttt.taxa <- pvclust(asv.depth.ttt.mean.pvclust, method.dist="cor", method.hclust="average", nboot=10000, parallel=TRUE)

plot(result.10k.ttt.taxa)
pvrect(result.10k.ttt.taxa, alpha=0.95)
```

## Toolik WS

Import mean abundance of taxa by depth for TWS heatmap
```{r}
asv.taxa.tws.mean <- read.csv("QIIME/SILVA/R_Data/taxa.mean.tws.csv")

colnames(asv.taxa.tws.mean)<-c("Phylum","0-10","10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90")

# Convert the first column (Phylum) into rownames
rownames(asv.taxa.tws.mean) <- asv.taxa.tws.mean$Phylum
asv.taxa.tws.mean<-as.data.frame(asv.taxa.tws.mean[-1])

# Convert dataframe into a matrix for heatmap
asv.taxa.tws.mean<-as.matrix(asv.taxa.tws.mean)

# Scale matrix values to generate Z-scores
asv.taxa.tws.mean<-scale(t(asv.taxa.tws.mean))

# Specify RColorBrewer custom color palette
col <- colorRampPalette(brewer.pal(10, "RdYlBu"))(256)
```

```{r}
# Pheatmap
asv.taxa.tws.mean.pheatmap <- pheatmap(asv.taxa.tws.mean, clustering_method = "average", cluster_cols = TRUE, cluster_rows = TRUE, cutree_cols = 3, cutree_rows = 3, angle_col=45, fontsize_col=8, legend = FALSE)

asv.taxa.tws.mean.pheatmap.v2 <- pheatmap(asv.taxa.tws.mean, clustering_method = "average", cluster_cols = TRUE, cluster_rows = FALSE, cutree_cols = 3, gaps_row = c(1, 5), angle_col=45, fontsize_col=10, legend = FALSE, cellheight=16, cellwidth=18)

#Export as .eps (width:650, height:600; "tws.heat.silva.eps")
```

### Soil Depth pvclust

TWS pvclust analysis
```{r}
asv.taxa.tws.mean.pvclust <- read.csv("QIIME/SILVA/R_Data/taxa.mean.tws.csv")

colnames(asv.taxa.tws.mean.pvclust)<-c("Phylum","0-10","10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90")

# Convert the first column (Phylum) into rownames
rownames(asv.taxa.tws.mean.pvclust) <- asv.taxa.tws.mean.pvclust$Phylum
asv.taxa.tws.mean.pvclust<-as.data.frame(asv.taxa.tws.mean.pvclust[-1])

# Scale data
asv.taxa.tws.mean.pvclust<-scale(asv.taxa.tws.mean.pvclust)
```

```{r}
result.10k.tws <- pvclust(asv.taxa.tws.mean.pvclust, method.dist="cor", method.hclust="average", nboot=10000, parallel=TRUE)

plot(result.10k.tws)
pvrect(result.10k.tws, alpha=0.95)
```

### Taxonomy pvclust

TWS Soil Taxa pvclust Analysis
```{r}
asv.depth.tws.mean.pvclust <- read.csv("QIIME/SILVA/R_Data/taxa.mean.tws.cols.csv")

# Convert the first column (Phylum) into rownames
rownames(asv.depth.tws.mean.pvclust) <- asv.depth.tws.mean.pvclust$Phylum
asv.depth.tws.mean.pvclust<-as.data.frame(asv.depth.tws.mean.pvclust[-1])

# Scale data
asv.depth.tws.mean.pvclust<-scale(asv.depth.tws.mean.pvclust)
```

```{r}
result.10k.tws.taxa <- pvclust(asv.depth.tws.mean.pvclust, method.dist="cor", method.hclust="average", nboot=10000, parallel=TRUE)

plot(result.10k.tws.taxa)
pvrect(result.10k.tws.taxa, alpha=0.95)
```

## Imnavait MAT

Import mean abundance of taxa by depth for ITT heatmap
```{r}
asv.taxa.itt.mean <- read.csv("QIIME/SILVA/R_Data/taxa.mean.itt.csv")

colnames(asv.taxa.itt.mean)<-c("Phylum","0-10","10-20","20-30","30-40","40-50","50-60","60-70")

# Convert the first column (Phylum) into rownames
rownames(asv.taxa.itt.mean) <- asv.taxa.itt.mean$Phylum
asv.taxa.itt.mean<-as.data.frame(asv.taxa.itt.mean[-1])

# Convert dataframe into a matrix for heatmap
asv.taxa.itt.mean<-as.matrix(asv.taxa.itt.mean)

# Scale matrix values to generate Z-scores
asv.taxa.itt.mean<-scale(t(asv.taxa.itt.mean))

# Specify RColorBrewer custom color palette
col <- colorRampPalette(brewer.pal(10, "RdYlBu"))(256)
```

```{r}
# Pheatmap
asv.taxa.itt.mean.pheatmap <- pheatmap(asv.taxa.itt.mean, clustering_method = "average", cluster_cols = TRUE, cluster_rows = TRUE, cutree_cols = 3, cutree_rows = 3, angle_col=45, fontsize_col=8, legend = FALSE)

asv.taxa.itt.mean.pheatmap.v2 <- pheatmap(asv.taxa.itt.mean, clustering_method = "average", cluster_cols = TRUE, cluster_rows = FALSE, cutree_cols = 3, gaps_row = c(2, 4), angle_col=45, fontsize_col=10, legend = FALSE, cellheight=16, cellwidth=17)

#Export as .eps (width:650, height:600; "itt.heat.silva.eps")
```

### Soil Depth pvclust

ITT pvclust analysis
```{r}
asv.taxa.itt.mean.pvclust <- read.csv("QIIME/SILVA/R_Data/taxa.mean.itt.csv")

colnames(asv.taxa.itt.mean.pvclust)<-c("Phylum","0-10","10-20","20-30","30-40","40-50","50-60","60-70")

# Convert the first column (Phylum) into rownames
rownames(asv.taxa.itt.mean.pvclust) <- asv.taxa.itt.mean.pvclust$Phylum
asv.taxa.itt.mean.pvclust<-as.data.frame(asv.taxa.itt.mean.pvclust[-1])

# Scale data
asv.taxa.itt.mean.pvclust<-scale(asv.taxa.itt.mean.pvclust)
```

```{r}
result.10k.itt <- pvclust(asv.taxa.itt.mean.pvclust, method.dist="cor", method.hclust="average", nboot=10000, parallel=TRUE)

plot(result.10k.itt)
pvrect(result.10k.itt, alpha=0.95)
```

### Taxonomy pvclust

ITT Soil Taxa pvclust Analysis
```{r}
asv.depth.itt.mean.pvclust <- read.csv("QIIME/SILVA/R_Data/taxa.mean.itt.cols.csv")

# Convert the first column (Phylum) into rownames
rownames(asv.depth.itt.mean.pvclust) <- asv.depth.itt.mean.pvclust$Phylum
asv.depth.itt.mean.pvclust<-as.data.frame(asv.depth.itt.mean.pvclust[-1])

# Scale data
asv.depth.itt.mean.pvclust<-scale(asv.depth.itt.mean.pvclust)
```

```{r}
result.10k.itt.taxa <- pvclust(asv.depth.itt.mean.pvclust, method.dist="cor", method.hclust="average", nboot=10000, parallel=TRUE)

plot(result.10k.itt.taxa)
pvrect(result.10k.itt.taxa, alpha=0.95)
```

## Imnavait WS

Import mean abundance of taxa by depth for IWS heatmap
```{r}
asv.taxa.iws.mean <- read.csv("QIIME/SILVA/R_Data/taxa.mean.iws.csv")

colnames(asv.taxa.iws.mean)<-c("Phylum","0-10","10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90","90-100")

# Convert the first column (Phylum) into rownames
rownames(asv.taxa.iws.mean) <- asv.taxa.iws.mean$Phylum
asv.taxa.iws.mean<-as.data.frame(asv.taxa.iws.mean[-1])

# Convert dataframe into a matrix for heatmap
asv.taxa.iws.mean<-as.matrix(asv.taxa.iws.mean)

# Scale matrix values to generate Z-scores
asv.taxa.iws.mean<-scale(t(asv.taxa.iws.mean))

# Specify RColorBrewer custom color palette
col <- colorRampPalette(brewer.pal(10, "RdYlBu"))(256)
```

```{r}
# Pheatmap
asv.taxa.iws.mean.pheatmap <- pheatmap(asv.taxa.iws.mean, clustering_method = "average", cluster_cols = TRUE, cluster_rows = TRUE, cutree_cols = 2, cutree_rows = 4, angle_col=45, fontsize_col=8, legend = FALSE)

asv.taxa.iws.mean.pheatmap.v2 <- pheatmap(asv.taxa.iws.mean, clustering_method = "average", cluster_cols = TRUE, cluster_rows = FALSE, cutree_cols = 3, gaps_row = c(1, 4, 6), angle_col=45, fontsize_col=10, legend = FALSE, cellheight=14, cellwidth=17)

#Export as .eps (width:650, height:600; "iws.heat.silva.eps")
```

### Soil Depth pvclust

IWS pvclust analysis
```{r}
asv.taxa.iws.mean.pvclust <- read.csv("QIIME/SILVA/R_Data/taxa.mean.iws.csv")

colnames(asv.taxa.iws.mean.pvclust)<-c("Phylum","0-10","10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90","90-100")

# Convert the first column (Phylum) into rownames
rownames(asv.taxa.iws.mean.pvclust) <- asv.taxa.iws.mean.pvclust$Phylum
asv.taxa.iws.mean.pvclust<-as.data.frame(asv.taxa.iws.mean.pvclust[-1])

# Scale data
asv.taxa.iws.mean.pvclust<-scale(asv.taxa.iws.mean.pvclust)
```

```{r}
result.10k.iws <- pvclust(asv.taxa.iws.mean.pvclust, method.dist="cor", method.hclust="average", nboot=10000, parallel=TRUE)

plot(result.10k.iws)
pvrect(result.10k.iws, alpha=0.9)
```

### Taxonomy pvclust

IWS Soil Taxa pvclust Analysis
```{r}
asv.depth.iws.mean.pvclust <- read.csv("QIIME/SILVA/R_Data/taxa.mean.iws.cols.csv")

# Convert the first column (Phylum) into rownames
rownames(asv.depth.iws.mean.pvclust) <- asv.depth.iws.mean.pvclust$Phylum
asv.depth.iws.mean.pvclust<-as.data.frame(asv.depth.iws.mean.pvclust[-1])

# Scale data
asv.depth.iws.mean.pvclust<-scale(asv.depth.iws.mean.pvclust)
```

```{r}
result.10k.iws.taxa <- pvclust(asv.depth.iws.mean.pvclust, method.dist="cor", method.hclust="average", nboot=10000, parallel=TRUE)
```

```{r}
plot(result.10k.iws.taxa)
pvrect(result.10k.iws.taxa, alpha=0.95)
```

## Sagwon MAT

Import mean abundance of taxa by depth for STT heatmap
```{r}
asv.taxa.stt.mean <- read.csv("QIIME/SILVA/R_Data/taxa.mean.stt.csv")

colnames(asv.taxa.stt.mean)<-c("Phylum","0-10","10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90")

# Convert the first column (Phylum) into rownames
rownames(asv.taxa.stt.mean) <- asv.taxa.stt.mean$Phylum
asv.taxa.stt.mean<-as.data.frame(asv.taxa.stt.mean[-1])

# Convert dataframe into a matrix for heatmap
asv.taxa.stt.mean<-as.matrix(asv.taxa.stt.mean)

# Scale matrix values to generate Z-scores
asv.taxa.stt.mean<-scale(t(asv.taxa.stt.mean))

# Specify RColorBrewer custom color palette
col <- colorRampPalette(brewer.pal(10, "RdYlBu"))(256)
```

```{r}
# Pheatmap
asv.taxa.stt.mean.pheatmap <- pheatmap(asv.taxa.stt.mean, clustering_method = "complete", cluster_cols = TRUE, cluster_rows = TRUE, cutree_cols = 2, cutree_rows = 2, angle_col=45, fontsize_col=8, legend = FALSE)

asv.taxa.stt.mean.pheatmap.v2 <- pheatmap(asv.taxa.stt.mean, clustering_method = "average", cluster_cols = TRUE, cluster_rows = FALSE, cutree_cols = 3, gaps_row = c(3,5), angle_col=45, fontsize_col=10, legend = FALSE, cellheight=16, cellwidth=18)

#Export as .eps (width:650, height:600; "stt.heat.silva.eps")
```

### Soil Depth pvclust

STT pvclust analysis
```{r}
asv.taxa.stt.mean.pvclust <- read.csv("QIIME/SILVA/R_Data/taxa.mean.stt.csv")

colnames(asv.taxa.stt.mean.pvclust)<-c("Phylum","0-10","10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90")

# Convert the first column (Phylum) into rownames
rownames(asv.taxa.stt.mean.pvclust) <- asv.taxa.stt.mean.pvclust$Phylum
asv.taxa.stt.mean.pvclust<-as.data.frame(asv.taxa.stt.mean.pvclust[-1])

# Scale data
asv.taxa.stt.mean.pvclust<-scale(asv.taxa.stt.mean.pvclust)
```

```{r}
result.10k.stt <- pvclust(asv.taxa.stt.mean.pvclust, method.dist="cor", method.hclust="average", nboot=10000, parallel=TRUE)

plot(result.10k.stt)
pvrect(result.10k.stt, alpha=0.95)
```

### Taxonomy pvclust

STT Soil Taxa pvclust Analysis
```{r}
asv.depth.stt.mean.pvclust <- read.csv("QIIME/SILVA/R_Data/taxa.mean.stt.cols.csv")

# Convert the first column (Phylum) into rownames
rownames(asv.depth.stt.mean.pvclust) <- asv.depth.stt.mean.pvclust$Phylum
asv.depth.stt.mean.pvclust<-as.data.frame(asv.depth.stt.mean.pvclust[-1])

# Scale data
asv.depth.stt.mean.pvclust<-scale(asv.depth.stt.mean.pvclust)
```

```{r}
result.10k.stt.taxa <- pvclust(asv.depth.stt.mean.pvclust, method.dist="cor", method.hclust="average", nboot=10000, parallel=TRUE)

plot(result.10k.stt.taxa)
pvrect(result.10k.stt.taxa, alpha=0.95)
```

## Sagwon WS

Import mean abundance of taxa by depth for SWS heatmap
```{r}
asv.taxa.sws.mean <- read.csv("QIIME/SILVA/R_Data/taxa.mean.sws.csv")

colnames(asv.taxa.sws.mean)<-c("Phylum","0-10","10-20","20-30","30-40","40-50","50-60","60-70")

# Convert the first column (Phylum) into rownames
rownames(asv.taxa.sws.mean) <- asv.taxa.sws.mean$Phylum
asv.taxa.sws.mean<-as.data.frame(asv.taxa.sws.mean[-1])

# Convert dataframe into a matrix for heatmap
asv.taxa.sws.mean<-as.matrix(asv.taxa.sws.mean)

# Scale matrix values to generate Z-scores
asv.taxa.sws.mean<-scale(t(asv.taxa.sws.mean))

# Specify RColorBrewer custom color palette
col <- colorRampPalette(brewer.pal(10, "RdYlBu"))(256)
```

```{r}
# Pheatmap
asv.taxa.sws.mean.pheatmap <- pheatmap(asv.taxa.sws.mean, clustering_method = "average", cluster_cols = TRUE, cluster_rows = TRUE, cutree_cols = 2, cutree_rows = 2, angle_col=45, fontsize_col=8, legend = FALSE)

asv.taxa.sws.mean.pheatmap.v2 <- pheatmap(asv.taxa.sws.mean, clustering_method = "average", cluster_cols = TRUE, cluster_rows = FALSE, cutree_cols = 2, gaps_row = 4, angle_col=45, fontsize_col=10, legend = FALSE, cellheight=16, cellwidth=16)

#Export as .eps (width:650, height:600; "sws.heat.silva.eps")
```

### Soil Depth pvclust

SWS pvclust analysis
```{r}
asv.taxa.sws.mean.pvclust <- read.csv("QIIME/SILVA/R_Data/taxa.mean.sws.csv")

colnames(asv.taxa.sws.mean.pvclust)<-c("Phylum","0-10","10-20","20-30","30-40","40-50","50-60","60-70")

# Convert the first column (Phylum) into rownames
rownames(asv.taxa.sws.mean.pvclust) <- asv.taxa.sws.mean.pvclust$Phylum
asv.taxa.sws.mean.pvclust<-as.data.frame(asv.taxa.sws.mean.pvclust[-1])

# Scale data
asv.taxa.sws.mean.pvclust<-scale(asv.taxa.sws.mean.pvclust)
```

```{r}
result.10k.sws <- pvclust(asv.taxa.sws.mean.pvclust, method.dist="cor", method.hclust="average", nboot=10000, parallel=TRUE)

plot(result.10k.sws)
pvrect(result.10k.sws, alpha=0.95)
```

### Taxonomy pvclust

SWS Soil Taxa pvclust Analysis
```{r}
asv.depth.sws.mean.pvclust <- read.csv("QIIME/SILVA/R_Data/taxa.mean.sws.cols.csv")

# Convert the first column (Phylum) into rownames
rownames(asv.depth.sws.mean.pvclust) <- asv.depth.sws.mean.pvclust$Phylum
asv.depth.sws.mean.pvclust<-as.data.frame(asv.depth.sws.mean.pvclust[-1])

# Scale data
asv.depth.sws.mean.pvclust<-scale(asv.depth.sws.mean.pvclust)
```

```{r}
result.10k.sws.taxa <- pvclust(asv.depth.sws.mean.pvclust, method.dist="cor", method.hclust="average", nboot=10000, parallel=TRUE)

plot(result.10k.sws.taxa)
pvrect(result.10k.sws.taxa, alpha=0.95)
```

```{r echo=FALSE, eval=FALSE}
# For calling plots and saving as .eps (width: 525; height: 650)
asv.taxa.ttt.mean.pheatmap.v2
asv.taxa.tws.mean.pheatmap.v2
asv.taxa.itt.mean.pheatmap.v2
asv.taxa.iws.mean.pheatmap.v2
asv.taxa.stt.mean.pheatmap.v2
asv.taxa.sws.mean.pheatmap.v2
```


# Reproducibility

The session information is provided for full reproducibility.
```{r}
devtools::session_info()
```
